---
title: "Introduction to SA24204150"
author: "Li Mengyao"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to SA24204150}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

<<<<<<< Updated upstream
**SA24204150** is an R package developed for the Statistical Computing course.
The package contain two parts:

1. Survival analysis and model evaluation, including three functions:
=======
**SA24204150** is an R package developed for the *Statistical Computing* course.
The package contain two parts:

1. **Survival analysis and model evaluation**, including three functions:
>>>>>>> Stashed changes

   * estimation of the censoring distribution,
   * Cox-model-based survival prediction,
   * IPCW Brier score evaluation;

<<<<<<< Updated upstream
2. **Algorithmic subset construction**, implemented by both R and C++.
=======
2. **Algorithmic subset construction**, implemented using both R and C++.
>>>>>>> Stashed changes

Specifically, the package contains three R functions for survival analysis,
namely *G.hat*, *S.hat*, and *bs*, and two Rcpp functions,
*generate_subsetsC* and *generate_nested_subsetsC*,
<<<<<<< Updated upstream
which are designed to generate subsets used in sequential modeling procedures.


## Subset construction using Rcpp

Subset generation is a fundamental building block in many statistical learning
and model selection algorithms.
In R, such operations are often implemented using high-level functions such as
*combn*, which can be computationally expensive for large input sizes.
In this package, two subset-generation functions are implemented in C++ using *Rcpp*.

### Generating all non-empty subsets

The Rcpp function *generate_subsetsC* generates all non-empty subsets of a given set.

The source C++ code is as follows:

List generate_subsetsC(IntegerVector set) {
int n = set.size();
int total = (1 << n);
List subsets;

for (int mask = 1; mask < total; ++mask) {
IntegerVector subset;
for (int i = 0; i < n; ++i) {
if (mask & (1 << i)) {
subset.push_back(set[i]);
}
}
subsets.push_back(subset);
}
return subsets;
}

This implementation enumerates subsets using bit masking, which is both concise
and efficient.

An example usage is shown below.

library(SA25204001)
generate_subsetsC(1:3)


### Generating nested subsets

In many greedy algorithms, only *nested* subsets are required.
The function *generate_nested_subsetsC* generates the sequence

[
{1}, {1,2}, \ldots, {1,\ldots,n}.
]

The corresponding C++ code is:

List generate_nested_subsetsC(IntegerVector set) {
int n = set.size();
List subsets(n);
for (int i = 0; i < n; ++i) {
subsets[i] = set[Range(0, i)];
}
return subsets;
}

An example is given below.

generate_nested_subsetsC(1:5)
=======
which are designed to generate subsets used in  sequential modeling procedures.

>>>>>>> Stashed changes


## Survival analysis functions

This package provides several R functions
for survival analysis and prediction.

### Estimation of the censoring survival function

The function *G.hat* computes the Kaplan–Meier estimator of the censoring survival
function, which is used for inverse probability of censoring weighting (IPCW).

The R source code is:

```{r, eval=FALSE}
G.hat <- function(t, Dat) {
newDat = list()
newDat$x = Dat$x
newDat$Delta = 1 - Dat$Delta
newDat$Y = Dat$Y
fit = survfit(Surv(Y, Delta) ~ 1, data = newDat)
if (t <= fit$time[1]) {
return(1)
} else {
q = max(which(fit$time <= t))
return(fit$surv[q])
}
}
```

An example of using *G.hat* is shown below.

```{r, eval=FALSE}
library(survival)
set.seed(123)
n <- 100
Dat <- list(
x = matrix(rnorm(n * 3), n, 3),
Y = rexp(n),
Delta = rbinom(n, 1, 0.7)
)
G.hat(0.5, Dat)
<<<<<<< Updated upstream

=======
```
>>>>>>> Stashed changes

### Cox-model-based survival prediction

The function *S.hat* fits a sequence of Cox proportional hazards models and
predicts survival probabilities at a fixed time point.

The R source code is:

```{r, eval=FALSE}
S.hat <- function(newX, newX_test, ns, ns_test, t, Datas, Datas_test, h) {
  p <- ncol(newX)
  K <- ceiling(p / h)
  
  S <- NULL
  S_test <- NULL
  
  Dat1 <- Datas
  Dat1$X <- newX  ## reordered/processed X
  
  Dat_test1 <- Datas_test
  Dat_test1$X <- newX_test
  
  pre.dat <- Dat1
  pre.dat$Y <- rep(t, ns)        ## fix time t
  
  pre.dat_test <- Dat_test1
  pre.dat_test$Y <- rep(t, ns_test)
  
  for (k in 1:K) {
    ind <- unique(c(1, ((k - 1) * h + 1):min(p, k * h)))
    res <- coxph(Surv(Y, as.numeric(Delta)) ~ X[, ind], data = Dat1)
    S.res <- stats::predict(res, newdata = pre.dat, type = "survival")
    S_test.res <- stats::predict(res, newdata = pre.dat_test, type = "survival")
    S <- cbind(S, S.res)
    S_test <- cbind(S_test, S_test.res)
  }
  
  list(S = S, S_test = S_test)
}
```

An example of using *S.hat* is shown below.

```{r, eval=FALSE}
h <- 2
newX <- Dat$x
newX_test <- matrix(rnorm(50 * 3), 50, 3)

Dat_test <- list(
x = newX_test,
Y = rexp(50),
Delta = rbinom(50, 1, 0.7)
)

res <- S.hat(
newX = newX,
newX_test = newX_test,
ns = n,
ns_test = 50,
t = 0.5,
Datas = Dat,
Datas_test = Dat_test
)
```

<<<<<<< Updated upstream
dim(res$S)

=======
>>>>>>> Stashed changes

### IPCW Brier score

Finally, the function *bs* computes the IPCW Brier score at a fixed time point.

The R source code is:

```{r, eval=FALSE}
bs <- function(newX, newX_test, ns, ns_test, t, Dat, Dat_test, k) {
  bs0 <- c()
  res.S <- S.hat(newX, newX_test, ns, ns_test, t, Dat, Dat_test)
  
  for (i in 1:ns) {
    if (Dat$Y[i] > t) {
      bs0 <- c(bs0, (1 - res.S$S[i, k])^2 / G.hat(t, Dat))
    } else if (as.numeric(Dat$Delta)[i] == 1) {
      bs0 <- c(bs0, (0 - res.S$S[i, k])^2 / G.hat(Dat$Y[i], Dat))
    }
  }
  
  sum(bs0) / ns
}
```

An example of using *bs* is shown below.

```{r, eval=FALSE}
bs(
newX = newX,
newX_test = newX_test,
ns = n,
ns_test = 50,
t = 0.5,
Dat = Dat,
Dat_test = Dat_test,
k = 1
)
```


## Subset construction using Rcpp

Subset generation is a fundamental building block in many statistical learning and model selection algorithms. In this package, two subset-generation functions are implemented in C++ using *Rcpp*.

### Generating all non-empty subsets

The Rcpp function *generate_subsetsC* generates all non-empty subsets of a given set.

The source C++ code is as follows:

```{r, eval=FALSE}
Rcpp::cppFunction(
List generate_subsetsC(IntegerVector set) {
int n = set.size();
int total = (1 << n);
List subsets;

for (int mask = 1; mask < total; ++mask) {
IntegerVector subset;
for (int i = 0; i < n; ++i) {
if (mask & (1 << i)) {
subset.push_back(set[i]);
}
}
subsets.push_back(subset);
}
return subsets;
}
)
```

This implementation enumerates subsets using bit masking, which is both concise and efficient.

An example usage is shown below.

```{r, eval=FALSE}
library(SA25204001)
generate_subsetsC(1:3)
```

### Generating nested subsets

In many greedy algorithms, only *nested* subsets are required.
The function *generate_nested_subsetsC* generates the sequence $\{1\}, \{1,2\}, \ldots, \{1,\ldots,n\}.$

The corresponding C++ code is:

```{r, eval=FALSE}
Rcpp::cppFunction(
List generate_nested_subsetsC(IntegerVector set) {
int n = set.size();
List subsets(n);
for (int i = 0; i < n; ++i) {
subsets[i] = set[Range(0, i)];
}
return subsets;
}
)
```

An example is given below.

```{r, eval=FALSE}
library(SA24204150)
generate_nested_subsetsC(1:5)
```

----

## Conclusion

This vignette demonstrates the usage of all major functions provided in the
package **SA25204001**.
The Rcpp functions offer efficient subset generation for algorithmic procedures and other R functions focus on survival analysis and model evaluation.


